Cultural Evolution of Conservation Behavior at the Intersection of Working

### 2020-10-06 DissertationProposal/DissertationProposal.Rmd
small changes.

# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common, however, for these areas to fail to reach their stated conservation goals. The most chronic causes of conservation project failure are the human dimensions. Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors. Research examining how and why conservation practices are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

@@

Globally, we have expended a tremendous amount of resources to identify conservation goals which will deliver the greatest benefits to biodiversity and ecosystem function. For these benefits to come to fruition however, conservation initiatives must be adopted at a rate which exceeds ecological degradation. This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.

@@

@@

```{r abmSim2, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under multiple group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmSim2.png")
```
# Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.

@@

Conservation via protected areas represents an ideal study system for testing the environmental applications of cultural group selection in the real world. Ignoring the regional to global scale ecosystem service benefits of protected areas, the units of the individual costs and group benefits of protected area adoption are directly comparable (i.e. fuelwood growth/use). This comparability will allow us to use established methods of measuring multilevel selection pressures to quantitatively predict if protected areas are likely to be selected for in each group. Finally, this project will compare the calculated selection pressures for protected area adoption from each group against the observed adoption of community managed protected areas.

### Question: 
Does cultural group selection theory accurately predict real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?

### Predictions
* P1: In wards where cultural group selection theory predicts that group selection should favor the adoption of community managed forested areas, we will observe significantly more real world adoption than in wards where adoption is not predicted to be favored by selection.

* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such as nonprofit conservation initiatives.

## Research Design
Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level. These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs). Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred. For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.

@@

```{r selPress, echo=FALSE, fig.cap="Simulated data showing protected area adoption vs calculated selection pressure for group adoption of protected areas.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/SelPressureAdopt.png")
```

@@

```{r csp2, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and geographic distance between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/CSP2.png")
```
# Conclusion
This research seeks to make clear how and why conservation behaviors succeed or fail to spread efficiently within a population. Moreover, the aim of this dissertation is to provide generalizable insights which can be applied to conservation initiatives in a variety of environments and population structures. A key advance made by this work is to link the fields of conservation social science and cultural group selection. Cultural group selection theory makes important predictions about how cooperative behaviors like conservation adoption may emerge and spread within populations, but has not yet been tested in a conservation context explicitly. Similarly, conservation social science incorporates environmental phenomena like resource mobility and regeneration, but largely omits generalizable insights about human behavioral evolution. By linking these two fields, we are greatly increasing the applicability of both bodies of work.

@@

>>>>

\newpage
# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common, however, for these areas to fail to reach their stated conservation goals. The most chronic causes of conservation project failure are the human dimensions. Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors. Research examining how and why conservation practices are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

@@

Globally, we have expended a tremendous amount of resources to identify conservation goals which will deliver the greatest benefits to biodiversity and ecosystem function. For these benefits to come to fruition however, conservation initiatives must be adopted at a rate which exceeds ecological degradation. This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.  
\newpage

@@

## Study System
All empirical research for this dissertation will take place on Pemba Island, Tanzania. In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar]. This population is divided into wards of 2,000 to 6,000 individuals. The majority of Pemban wards are described as forest-dependent. The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, established forest protected areas and exclude outside individuals [@benjaminsen2014between]. As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].

\newpage

@@

```{r abmSim2, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under multiple group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmSim2.png")
```
\newpage

# Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.

@@

Protected areas are an ideal study system for testing the mechanisms of cultural group selection in real world conservation. When a community adopts a protected area, like the Community Forest Management Agreements observed on Pemba Island, Tanzania, individuals agree to forego a portion of their personal harvest potential for the benefit of the community. Under cultural group selection, we only expect this behavior to emerge when the proportion of dissimilar to similar cultural traits between a group and its neighbors outweighs the benefit/cost ratio of engaging in cooperative conservation. Plainly put, communities should be more likely to adopt CoFMAs if they are culturally dissimilar to their surrounding communities. This project will test this theory by evaluating the perceived individual costs and group benefits of CoFMAs, within and between group cultural similarities (cultural $F_{st}$), and actual adoption of CoFMAs in forest dependent wards on Pemba Island [@kline2018designing].

### Question: 
Do Pemban wards with stronger cultural differentiation from surrounding wards adopt CoFMAs at a higher rate than wards who are more culturally similar to their neighbors?

### Predictions
* P1: Cultural differentiation of Pemban wards will predict CoFMA adoption only when the perceived individual costs of CoFMAs are relatively low.

* P2: High levels of forest dependence and perceived costs of CoFMAs will prevent CoFMA adoption, even when groups are considerably culturally dissimilar to adjacent groups.

## Research Design
Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level. These group level selection pressures are what we hypothesize to drive the adoption of CoFMAs. Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred. For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.

@@

```{r selPress, echo=FALSE, fig.cap="Simulated data showing protected area adoption vs calculated selection pressure for group adoption of protected areas.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/SelPressureAdopt.png")
```
\newpage

@@

```{r csp2, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and geographic distance between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/CSP2.png")
```
\newpage

# Conclusion
This research seeks to make clear how and why conservation behaviors succeed or fail to spread efficiently within a population. Moreover, the aim of this dissertation is to provide generalizable insights which can be applied to conservation initiatives in a variety of environments and population structures. A key advance made by this work is to link the fields of conservation social science and cultural group selection. Cultural group selection theory makes important predictions about how cooperative behaviors like conservation adoption may emerge and spread within populations, but has not yet been tested in a conservation context explicitly. Similarly, conservation social science incorporates environmental phenomena like resource mobility and regeneration, but largely omits generalizable insights about human behavioral evolution. By linking these two fields, we are greatly increasing the applicability of both bodies of work.

@@

\newpage
